Estimation below the detection limit

### 2018-04-13 TEX/draft2.tex
ran an experiment to check if optimization is slow because of gammas: no, it's just slow.
Launching a big XP

For the 1D experiment, we fix $K = 3$ signals of length $L = 21$, as depicted in Figure~\TODO{ref}. Following the data model described in Section~\TODO{ref}, we generate an observation $y$ of length $24.6 \cdot 10^9$. Each of the three signals appears, respectively (and approximately) $30 \cdot 10^6$, $20 \cdot 10^6$ and $10 \cdot 10^6$ times in $y$, such that at least $L-1$ zeros separate two occurrences of any signals. This is done by randomly selecting $60 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint and locations picked so far. For each placement, one of the three signals is picked at random proportionally to the desired number of occurrences of each. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The SNR of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/12.
%Given the scaling of the target signals (whose median absolute value for the entries is at or below one),
This is enough noise to make cross-correlations of $y$ even with the true signals display peaks at random locations, uninformative of the actual locations of the signal occurrences. Thus, we contend that it would be difficult for any algorithm to locate the signal occurrences, let alone to classify them according to which signal appears where.

@@

Having computed the moments of interest, we now estimate signals $x_1, \ldots, x_K$ and coefficients $\gamma_1, \ldots, \gamma_K$ which agree with the data. We choose to do so by running an optimization algorithm on the following nonlinear least-squares problem:
\begin{multline}
	\min_{\substack{\hat x_1, \ldots, \hat x_K \in \reals^{W} \\ \hat \gamma_1, \ldots, \hat \gamma_K > 0}} w_1 \left( a_y^1 - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^1 \right)^2 + w_2 \sum_{\ell = 1}^{L-1} \left( a_y^2[\ell] - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^2[\ell] \right)^2 + \\ w_3 \sum_{\substack{2\leq\ell_1\leq L-1 \\ 1 \leq \ell_2 \leq \ell_1-1}} \left( a_y^3[\ell_1, \ell_2] - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^3[\ell_1,\ell_2] \right)^2.
	\label{eq:optim1D}
\end{multline}
where $W \geq L$ is the length of the sought signals and \TODO{explain $w_i$'s: currently they are $w_1 = 1/2, w_2 = 1/2n_2, w_3 = 1/2n_3$, where $n_2, n_3$ are the number of moments used: $n_2 = L-1$, $n_3 = \frac{(L-1)(L-2)}{2}$. Issue is: this is not very smart..}. Setting $W = L$ (as is a priori desired) is problematic because the above optimization problems appears to have numerous poor local optimizers.
%Since we can only initialize randomly at first, this approach would often fail in practice. Alternatively,
Thus, we first run the optimization with $W = 2L-1$. This problem appears to have fewer poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded to length $W$, we extract from each one a subsignal of length $L$ (with cyclic indexing \TODO{we should understand / explain this}) that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt}, which allows to treat the positivity constraints \TODO{I might need a reference for this} on coefficients $\hat \gamma_k$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.
\TODO{Should we do variable projection for the gammas, that is, exploit the fact the problem is a regular least squares in the gammas (up to the positivity constraints) to substitute the explicit optimum for them? Not sure it's worth the effort.}

>>>>

For the 1D experiment, we fix $K = 3$ signals of length $L = 21$, as depicted in Figure~\TODO{ref}. Following the data model described in Section~\TODO{ref}, we generate an observation $y$ of length $24.6 \cdot 10^9$. Each of the three signals appears, respectively (and approximately) $300 \cdot 10^6$, $200 \cdot 10^6$ and $100 \cdot 10^6$ times in $y$, such that at least $L-1$ zeros separate two occurrences of any signals. This is done by randomly selecting $600 \cdot 10^6$ placements in $y$, one at a time with an accept/reject rule based on the separation constraint and locations picked so far. For each placement, one of the three signals is picked at random proportionally to the desired number of occurrences of each. Then, i.i.d.\ Gaussian noise with mean zero and standard deviation $\sigma = 3$ is added, to form the observed $y$. The SNR of $y$
% sqrt((m_want*sum(X.^2)')/(sigma^2*n))
is about 1/12.
%Given the scaling of the target signals (whose median absolute value for the entries is at or below one),
This is enough noise to make cross-correlations of $y$ even with the true signals display peaks at random locations, uninformative of the actual locations of the signal occurrences. Thus, we contend that it would be difficult for any algorithm to locate the signal occurrences, let alone to classify them according to which signal appears where.

@@

Having computed the moments of interest, we now estimate signals $x_1, \ldots, x_K$ and coefficients $\gamma_1, \ldots, \gamma_K$ which agree with the data. We choose to do so by running an optimization algorithm on the following nonlinear least-squares problem:
\begin{multline}
	\min_{\substack{\hat x_1, \ldots, \hat x_K \in \reals^{W} \\ \hat \gamma_1, \ldots, \hat \gamma_K > 0}} w_1 \left( a_y^1 - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^1 \right)^2 + w_2 \sum_{\ell = 1}^{L-1} \left( a_y^2[\ell] - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^2[\ell] \right)^2 + \\ w_3 \sum_{\substack{2\leq\ell_1\leq L-1 \\ 1 \leq \ell_2 \leq \ell_1-1}} \left( a_y^3[\ell_1, \ell_2] - \sum_{k=1}^K \hat \gamma_k a_{\hat x_k}^3[\ell_1,\ell_2] \right)^2.
	\label{eq:optim1D}
\end{multline}
where $W \geq L$ is the length of the sought signals and \TODO{explain $w_i$'s: currently they are $w_1 = 1/2, w_2 = 1/2n_2, w_3 = 1/2n_3$, where $n_2, n_3$ are the number of moments used: $n_2 = L-1$, $n_3 = \frac{(L-1)(L-2)}{2}$. Issue is: this is not very smart..}. Setting $W = L$ (as is a priori desired) is problematic because the above optimization problems appears to have numerous poor local optimizers.
%Since we can only initialize randomly at first, this approach would often fail in practice. Alternatively,
Thus, we first run the optimization with $W = 2L-1$. This problem appears to have fewer poor local optima, perhaps because the additional degrees of freedom allow for more escape directions. Since we hope the signals estimated this way correspond to the true signals zero-padded to length $W$, we extract from each one a subsignal of length $L$ (with cyclic indexing \TODO{we should understand / explain this}) that has largest $\ell_2$-norm. This estimator is then used as initial iterate for~\eqref{eq:optim1D}, this time with $W = L$. We find that this procedure is reliable for a wide range of experimental parameters. To solve~\eqref{eq:optim1D}, we run the trust-region method implemented in Manopt~\cite{manopt}, which allows to treat the positivity constraints \TODO{I might need a reference for this} on coefficients $\hat \gamma_k$. Notice that the cost function is a polynomial in the variables, so that it is straightforward to compute it and its derivatives.
\TODO{Should we do variable projection for the gammas, that is, exploit the fact the problem is a regular least squares in the gammas (up to the positivity constraints) to substitute the explicit optimum for them? Not sure it's worth the effort. -- Ok, it's probably not a good idea, because even with fixed gammas to the correct value, optimization takes a while.}
\TODO{Do we still need to stress at this point that the optimization part has complexity independent of length of observation? Should be pretty clear at this point already.}
